### 2020-10-13 Analysis/Code/Sorokin_JMP_2020.Rmd
misc

@SorokinStevens20 raises other concerns about the validity of the regression discontinuity approach in studying the causal effect of reviews on product outcomes. To my knowledge, it was the first paper to use Steam to study the causal effect of online reviews. In the absence of direct sales data, @SorokinStevens20 uses a particular regression specification that extracts the information on sales from video games' usage data. I inherit the data from that paper, but confront the absence of sales data in a new way. I make assumptions about the gaming patterns for the games in my sample, and formulate a structural model of the data generating process. With the help of this model, I manage to extract information on product sales from usage data. I compare the estimates of the structural model with the ones obtained by its simpler regression analog, and find important similartities and dissimilarities. The last improvement of the present study over @SorokinStevens20 is that I employ all the relevant review tiers present on Steam, while in earlier work the regression discontinuity approach dictated the focus on a certain subset of review labels.

@@

# Arrival of reviews relative to the number accumulated at 180 days, by quintiles
# Assign a number to each game that measures the number of full months it has
# spent in the sample
panel[, ageGroup:=age%/%30][, ageGroup:=rep(max(ageGroup), .N), by = ID]
M <- panel[, max(ageGroup)]
# Create an auxilary table that for every age (in months) takes games that have
# been present for that many months, and then splits those games into 10 groups,
# based on the amount of reviews they had at that age.
f <- data.table(age=integer(), sizeGroup=integer(), ID=integer())
for(m in 1:M){
  f<-rbindlist(list(f,
                    panel[ageGroup>=m&age==30*m, 
                          .(age, sizeGroup=ecdf(reviews)(reviews), ID)
                          ][, sizeGroup:=cut(sizeGroup,nGroups,labels=seq(1:nGroups))]),
               use.names=TRUE)
}
setkey(f, age, sizeGroup, ID)

@@

Before I start my investigation into the discounting behavior of the products close to a review transition, it is important to address two question. First, do discounts matter on Steam? Do they generate enough additional sales to be relevant? While the answer might be obvious for large sought-after products, the situation is less clear for an average game. The second question is: how big are the differences between various review tiers? If transitioning to a better tier does not increase sales by much, then firms would be unlikely to pursue such transitions when making their discount decisions. The discussion of the assumptions behind my analysis and the interpretation of the parameter estimates contributes greatly to the size of this section. However, the main takeaways are simple: better review labels improve sales, and discounts are important drivers of sales on Steam. Having few reviews or having negative reviews depresses sales by 5-13% compared to having “Mixed” reviews; the “Overwhelmingly Positive” review tier increases sales by about 10% compared to “Mixed”. An average discount in the sample increases sales by 38% upon introduction, and by another 3-12% every day thereafter. These two facts form the foundation of my analysis of discounting behavior on a verge of a review tier transition that will be carried out in Section \ref{descriptiveEvidence}. I will further extend the model from this section by modelling the reviewing behavior of customers in Section \ref{selectionEffect}.

@@

There are two key identification challenges I need to address in order to estimate the parameters of the demand model from my data. One challenge comes from the fact that in order to tease out new purchases from the player activity data, I need to know something about player exit, which is also unobserved. The problem is illustrated by equation (\ref{activeProccess}), where the observed part of the data $(A_{it} - A_{it-1})$ only identifies the difference $(B_{it} - E_{it})$. Consider two  hypothetical games that have the same observed median daily player count---say, ten people. However, game one offers a lot of replayability and is played by the same ten people over and over again, while game two is played by new ten people every day. Clearly, such two products with identical player activity patterns would have very different sales. I am able to solve this identification problem with the help of Assumption \ref{arrivalAssumptions}. The key insight is that observed factors determine the arrival of new players (buyers) in *absolute* terms, while exit is *proportional* to the number of active players. Intuitively, the exit procees is identified by the rate of decay of the player count when the count is far from its (slowly-changing) trend or average. Player count is typically far from its average on the days following the game release or the introduction of a discount. On such occasions a lot of new players buy the game simultaneously, which leads to sharp spikes in the player count variable (see example game in Figure \ref{exampleGame}). For a given decay parameter, the level at which the player count settles identifies the flow rate of new buyers. Given that identifying the exit rate is crucial for teasing out the sales component, it is a great advantage of my model that it allows full heterogeneity in the exit rate $(1-\psi_i)$.

@@

The substitution of the week dummies with a Seasonal Sale dummy proved to be sufficient for convergence. I mentioned in the Identification section that a more simplistic, yet more tractable, alternative to estimating the non-linear model (\ref{arrivalReg}) is given by equation (\ref{logReg})
$$
\log A_{it} = \tilde\lambda_i + \tilde\psi_i \log A_{it-1} + x'_{it}\tilde \beta + u_{it}
$$
\noindent To check whether the omitted week effects could play a crucial role, I report the results of estimaing (\ref{logReg}) both with the ideal set of covariates, and with the covariates used to estimate (\ref{arrivalReg}). The results are presented in Table \ref{demandEstimates}.

@@

However, columns (1) and (2)-(3) differ substantially in the estimates of the effect of having no score on sales\footnote{In order to have both the score variable $\in [0,100]$ and a dummy for the “No Score” label I set the former to be equal to 0 when a game has no review label (which happens when it has less than 10 reviews). Thus, the score variable really measures the effect of score once the score is defined.}. While the structural model (\ref{arrivalReg}) suggests that having no score is associated with a `r -round(beta.hat[6]*100)`% slower customer arrival than having the “Mixed” review label, its simplified log-regression analogs find the opposite effect of a having `r round(reg1$coefficients[5]*100)`-`r round(reg2$coefficients[4]*100)`% bonus associated with having not enough reviews. The latter effect can not possibly be taken at face value, given that the most exclusive review label, "Overwhelmingly Positive",  increases sales by at most `r round(beta.hat[11]*100)`% compared to the baseline. It would be possible to explain such large estimates by high levels of demand  when the game is young and is likely to have less than 10 reviews, but I control for being “young” by including the dummy for being less than 14 days of age\footnote{A two weeks cutoff is inspired by high speed of review arrival documented in Figure \ref{reviewDescriptivePlots}}. Thus, the coefficient on “No Score” is identified by games that exit this bin later than their first two weeks. Such games constitute `r round(100*panel[noScore==T & age>=14, length(unique(ID))]/info[,.N])`% of the sample, and are probably different from the rest of the games. However, it is still hard to come up with omitted factors that would explain the positive effect of “No Score” on sales found in columns (2)-(3). One explanation that works is that the log-regression simplification (\ref{logReg}) is simply misguided, and that one should only trust the coefficients from the structural model of the demand process (\ref{arrivalReg}). This latter model finds a penalty of `r -round(beta.hat[6]*100)`% associated with  having no score, which sounds much more reasonable. A negative effect could be attributed to customers' reluctance to purchase products of unknown quality. If that is the case, this finding suggests a pretty serious cost of asymetric information in this market.

The second imprortant set of variables are the price variables. The first row of Table \ref{demandEstimates} suggests that the price elasticity of demand of very low, between  `r round(-reg1$coefficients[2],2)` and `r round(-beta.hat[1],2)` in absolute value. This magnitude is very similar to the price elasticity that @ReimersWaldfogel20 finds studying reviews for books on Amazon.com. That paper argues that Amazon has been known to prioritize growth over profit, thus charging low prices and operating on the inelastic part of the demand curve. However, on Steam individual firms choose which prices to charge for their product, and they should have a stronger preference for profit than an entity that is a marketplace, rather than a seller. The situation becomes more clear after examining the coefficient on “New Discount”, an indicator variable for the first day after the introduction of the discount\footnote{I found the effect to be stronger if a one day lag is allowed}. As I have mentioned in the Identification section, the durable good nature of video games means that forward-looking behavior by consumers could lead to different price elasticities for short-lived promotions and long-term price changes, and the difference between the price coefficient and the indicator for the start of a discount could capture precisely that distinction. I find that, on average, a discount leads to a `r round(beta.hat[2]*100)`-`r round(reg2$coefficients[3]*100)`% spike in sales upon introduction. An average discount in the sample is `r round(panel[discount>0, mean(discount)]*100)`%, meaning that, based on the estimates in column 1, the effective change in quantity sold upon the introduction of the discount is `r round(beta.hat[2]*100)`% $+$ `r round(-beta.hat[1]*round(panel[discount>0, mean(discount)]*100))`% $=$ `r round(beta.hat[2]*100) + round(-beta.hat[1]*round(panel[discount>0, mean(discount)]*100))`%, with the implied elasticity of `r round((round(beta.hat[2]*100) + round(-beta.hat[1]*round(panel[discount>0, mean(discount)]*100)))/round(panel[discount>0, mean(discount)]*100),2)`. This number looks more reasonable, especially given that the products I study have a zero marginal cost of production, and in a static world firms would set prices at which the demand for their products would be unit-elastic. Every additional day on sale  is then associated with a `r round(round(-reg1$coefficients[2],2)*round(panel[discount>0, mean(discount)]*100))`-`r round(round(-beta.hat[1],2)*round(panel[discount>0, mean(discount)]*100))`% higher quantity demanded, where larger effect comes from the structural model. Discounts on Steam attract a considerable number of new users, and thus could be effective tools in generating new reviews needed for a review tier transition.

@@

I estimate (\ref{potTransitionsHeteroReg}) by OLS (without the fixed effect term $f_i$) and by the within-estimator. The results are presented in Table \ref{scoreDifTransDiscReg} (the controls are omitted for brevity). The least squares estimates in column (OLS) are mainly presented to check the correlations in the data. Games that tend to receive better reviews on discounts discount more. However, the selection effect is not significant in both specifications. Crucially, in the preferred specification (FE) I find that $\beta^-_0 < 0$, and $\beta^+_0 > 0$. Once the variation in the selection effect is controlled for, I obtain the coefficients measuring the impact of the variance effect on discounting decisions, and the signs align with theory. Their sizes are also quite economically meaningful, suggesting that, in the absence of the selection effect, games close to a positive transition are  `r round(100*reg3$coefficients["pos.treatmentTRUE"]/panel[, mean((discount > 0), na.rm=T)])`% more likely to go on a discount, and games close to a negative transition are `r -round(100*reg3$coefficients["neg.treatmentTRUE"]/panel[, mean((discount > 0), na.rm=T)])`% less likely to go on a discount. Recall that my estimates of (\ref{potTransitionsReg}) that did not control for the selection effect found any effect of the proximity to a negative transition on the probability of discount.

@@

In this paper I document a novel mechanism through which firms use pricing to manage their online reviews. Using data from Steam, a major digital marketplace for computer video games, I show that firms discount their products in order to facilitate transitions between review tiers. My estimates of the demand process indicate that better review tiers substantially increase sales, and that discounts are effective at attracting new users and, therefore, reviewers. Therefore, using price promotions to improve the review label of a game appears to be a plausible strategy. I find that products that are close to upgrading their review tier are 4-9% more likely to be on a discount, while products close to deteriorating their review tier are up to 4% less likely to discount, albeit this effect is less significant.

I identify two channels that can make such behavior profitable: the *selection* effect and the *variance* effect. Customers who purchase the good on discount are likely to be different from the ones who pay the full price, and could potentially leave different reviews. If such reviews tend to be more favorable, then the induced positive selection effect creates incentives for the product to go on discount, regardless of whether the product is about to upgrade or downgrade its review tier. There is some disagreement in the literature about the sign of the selection effect. I contribute by showing that on Steam the effect is not positive.

The variance effect reflects the asymmetry between a successful and an unsuccessful discounting campaign, as measured by the valence of the new reviews. A product close to moving one review tier up benefits greatly from a good realization, while a bad might not change its review tier. On the other hand, a product that is close to sliding into a worse tier is averse to a bad outcome, and would benefit less in absolute value from a good outcome. While I manage to test the selection effect directly, I have to test the variance effect by trying to falsify the predictions that follow from it. My strongest test is given by the prediction that, in the absence of a positive selection effect, the variance effect has to make a product on the verge of deteriorating its reviews less likely to discount. When I control for the size of the selection effect, I find that the aforementioned decline in probability of discounting goes from being 4% and weakly significant to 6% and highly significant. I also leverage my unique data and  structural model to document that consumers are more likely to leave reviews when they buy on a discount, which indirectly supports the importance of the variance effect.

>>>>

@SorokinStevens20 raises other concerns about the validity of the regression discontinuity approach in studying the causal effect of reviews on product outcomes. To my knowledge, it was the first paper to use Steam to study the causal effect of online reviews. In the absence of direct sales data, @SorokinStevens20 uses a particular regression specification that extracts the information on sales from video games' usage data. I inherit the data from that paper, but confront the absence of sales data in a new way. I make assumptions about the gaming patterns for the games in my sample, and formulate a structural model of the data generating process. With the help of this model, I manage to extract information on product sales from usage data. I compare the estimates of the structural model with the ones obtained by its simpler regression analog, and find important similarities and dissimilarities. The last improvement of the present study over @SorokinStevens20 is that I employ all the relevant review tiers present on Steam, while in earlier work the regression discontinuity approach dictated the focus on a certain subset of review labels.

@@

# Arrival of reviews relative to the number accumulated at 180 days, by quintiles
# Assign a number to each game that measures the number of full months it has
# spent in the sample
panel[, ageGroup:=age%/%30][, ageGroup:=rep(max(ageGroup), .N), by = ID]
M <- panel[, max(ageGroup)]
# Create an auxiliary table that for every age (in months) takes games that have
# been present for that many months, and then splits those games into 10 groups,
# based on the amount of reviews they had at that age.
f <- data.table(age=integer(), sizeGroup=integer(), ID=integer())
for(m in 1:M){
  f<-rbindlist(list(f,
                    panel[ageGroup>=m&age==30*m, 
                          .(age, sizeGroup=ecdf(reviews)(reviews), ID)
                          ][, sizeGroup:=cut(sizeGroup,nGroups,labels=seq(1:nGroups))]),
               use.names=TRUE)
}
setkey(f, age, sizeGroup, ID)

@@

Before I start my investigation into the discounting behavior of the products close to a review transition, it is important to address two question. First, do discounts matter on Steam? Do they generate enough additional sales to be relevant? While the answer might be obvious for large sought-after products, the situation is less clear for an average game. The second question is: how big are the differences between various review tiers? If transitioning to a better tier does not increase sales by much, then firms would be unlikely to pursue such transitions when making their discount decisions. The discussion of the assumptions behind my analysis and the interpretation of the parameter estimates contributes greatly to the size of this section. However, the main takeaways are simple: better review labels improve sales, and discounts are important drivers of sales on Steam. Having few reviews or having negative reviews depresses sales by 5-13% compared to having “Mixed” reviews; the “Overwhelmingly Positive” review tier increases sales by about 10% compared to “Mixed”. An average discount in the sample increases sales by 38% upon introduction, and by another 3-12% every day thereafter. These two facts form the foundation of my analysis of discounting behavior on a verge of a review tier transition that will be carried out in Section \ref{descriptiveEvidence}. I will further extend the model from this section by modeling the reviewing behavior of customers in Section \ref{selectionEffect}.

@@

There are two key identification challenges I need to address in order to estimate the parameters of the demand model from my data. One challenge comes from the fact that in order to tease out new purchases from the player activity data, I need to know something about player exit, which is also unobserved. The problem is illustrated by equation (\ref{activeProccess}), where the observed part of the data $(A_{it} - A_{it-1})$ only identifies the difference $(B_{it} - E_{it})$. Consider two  hypothetical games that have the same observed median daily player count---say, ten people. However, game one offers a lot of replayability and is played by the same ten people over and over again, while game two is played by new ten people every day. Clearly, such two products with identical player activity patterns would have very different sales. I am able to solve this identification problem with the help of Assumption \ref{arrivalAssumptions}. The key insight is that observed factors determine the arrival of new players (buyers) in *absolute* terms, while exit is *proportional* to the number of active players. Intuitively, the exit process is identified by the rate of decay of the player count when the count is far from its (slowly-changing) trend or average. Player count is typically far from its average on the days following the game release or the introduction of a discount. On such occasions a lot of new players buy the game simultaneously, which leads to sharp spikes in the player count variable (see example game in Figure \ref{exampleGame}). For a given decay parameter, the level at which the player count settles identifies the flow rate of new buyers. Given that identifying the exit rate is crucial for teasing out the sales component, it is a great advantage of my model that it allows full heterogeneity in the exit rate $(1-\psi_i)$.

@@

The substitution of the week dummies with a Seasonal Sale dummy proved to be sufficient for convergence. I mentioned in the Identification section that a more simplistic, yet more tractable, alternative to estimating the non-linear model (\ref{arrivalReg}) is given by equation (\ref{logReg})
$$
\log A_{it} = \tilde\lambda_i + \tilde\psi_i \log A_{it-1} + x'_{it}\tilde \beta + u_{it}
$$
\noindent To check whether the omitted week effects could play a crucial role, I report the results of estimating (\ref{logReg}) both with the ideal set of covariates, and with the covariates used to estimate (\ref{arrivalReg}). The results are presented in Table \ref{demandEstimates}.

@@

However, columns (1) and (2)-(3) differ substantially in the estimates of the effect of having no score on sales\footnote{In order to have both the score variable $\in [0,100]$ and a dummy for the “No Score” label I set the former to be equal to 0 when a game has no review label (which happens when it has less than 10 reviews). Thus, the score variable really measures the effect of score once the score is defined.}. While the structural model (\ref{arrivalReg}) suggests that having no score is associated with a `r -round(beta.hat[6]*100)`% slower customer arrival than having the “Mixed” review label, its simplified log-regression analogs find the opposite effect of a having `r round(reg1$coefficients[5]*100)`-`r round(reg2$coefficients[4]*100)`% bonus associated with having not enough reviews. The latter effect can not possibly be taken at face value, given that the most exclusive review label, "Overwhelmingly Positive",  increases sales by at most `r round(beta.hat[11]*100)`% compared to the baseline. It would be possible to explain such large estimates by high levels of demand  when the game is young and is likely to have less than 10 reviews, but I control for being “young” by including the dummy for being less than 14 days of age\footnote{A two weeks cutoff is inspired by high speed of review arrival documented in Figure \ref{reviewDescriptivePlots}}. Thus, the coefficient on “No Score” is identified by games that exit this bin later than their first two weeks. Such games constitute `r round(100*panel[noScore==T & age>=14, length(unique(ID))]/info[,.N])`% of the sample, and are probably different from the rest of the games. However, it is still hard to come up with omitted factors that would explain the positive effect of “No Score” on sales found in columns (2)-(3). One explanation that works is that the log-regression simplification (\ref{logReg}) is simply misguided, and that one should only trust the coefficients from the structural model of the demand process (\ref{arrivalReg}). This latter model finds a penalty of `r -round(beta.hat[6]*100)`% associated with  having no score, which sounds much more reasonable. A negative effect could be attributed to customers' reluctance to purchase products of unknown quality. If that is the case, this finding suggests a pretty serious cost of asymmetric information in this market.

The second important set of variables are the price variables. The first row of Table \ref{demandEstimates} suggests that the price elasticity of demand of very low, between  `r round(-reg1$coefficients[2],2)` and `r round(-beta.hat[1],2)` in absolute value. This magnitude is very similar to the price elasticity that @ReimersWaldfogel20 finds studying reviews for books on Amazon.com. That paper argues that Amazon has been known to prioritize growth over profit, thus charging low prices and operating on the inelastic part of the demand curve. However, on Steam individual firms choose which prices to charge for their product, and they should have a stronger preference for profit than an entity that is a marketplace, rather than a seller. The situation becomes more clear after examining the coefficient on “New Discount”, an indicator variable for the first day after the introduction of the discount\footnote{I found the effect to be stronger if a one day lag is allowed}. As I have mentioned in the Identification section, the durable good nature of video games means that forward-looking behavior by consumers could lead to different price elasticities for short-lived promotions and long-term price changes, and the difference between the price coefficient and the indicator for the start of a discount could capture precisely that distinction. I find that, on average, a discount leads to a `r round(beta.hat[2]*100)`-`r round(reg2$coefficients[3]*100)`% spike in sales upon introduction. An average discount in the sample is `r round(panel[discount>0, mean(discount)]*100)`%, meaning that, based on the estimates in column 1, the effective change in quantity sold upon the introduction of the discount is `r round(beta.hat[2]*100)`% $+$ `r round(-beta.hat[1]*round(panel[discount>0, mean(discount)]*100))`% $=$ `r round(beta.hat[2]*100) + round(-beta.hat[1]*round(panel[discount>0, mean(discount)]*100))`%, with the implied elasticity of `r round((round(beta.hat[2]*100) + round(-beta.hat[1]*round(panel[discount>0, mean(discount)]*100)))/round(panel[discount>0, mean(discount)]*100),2)`. This number looks more reasonable, especially given that the products I study have a zero marginal cost of production, and in a static world firms would set prices at which the demand for their products would be unit-elastic. Every additional day on sale  is then associated with a `r round(round(-reg1$coefficients[2],2)*round(panel[discount>0, mean(discount)]*100))`-`r round(round(-beta.hat[1],2)*round(panel[discount>0, mean(discount)]*100))`% higher quantity demanded, where larger effect comes from the structural model. Discounts on Steam attract a considerable number of new users, and thus could be effective tools in generating new reviews needed for a review tier transition.

@@

I estimate (\ref{potTransitionsHeteroReg}) by OLS (without the fixed effect term $f_i$) and by the within-estimator. The results are presented in Table \ref{scoreDifTransDiscReg} (the controls are omitted for brevity). The least squares estimates in column (OLS) are mainly presented to check the correlations in the data. Games that tend to receive better reviews on discounts discount more. However, the selection effect is not significant in both specifications. Crucially, in the preferred specification (FE) I find that $\beta^-_0 < 0$, and $\beta^+_0 > 0$. Once the variation in the selection effect is controlled for, I obtain the coefficients measuring the impact of the variance effect on discounting decisions, and the signs align with theory. Their sizes are also quite economically meaningful, suggesting that, in the absence of the selection effect, games close to a positive transition are  `r round(100*reg3$coefficients["pos.treatmentTRUE"]/panel[, mean((discount > 0), na.rm=T)])`% more likely to go on a discount, and games close to a negative transition are `r -round(100*reg3$coefficients["neg.treatmentTRUE"]/panel[, mean((discount > 0), na.rm=T)])`% less likely to go on a discount. Recall that my estimates of (\ref{potTransitionsReg}) that did not control for the selection effect, found only a limited effect of the proximity to a negative transition on the probability of discount.

@@

In this paper I document a novel mechanism through which firms use pricing to manage their online reviews. Using data from Steam, a major digital marketplace for computer video games, I show that firms discount their products in order to facilitate transitions between review tiers. My estimates of the demand process indicate that better review tiers substantially increase sales, and that discounts are effective at attracting new users and, as a consequence, reviewers. Therefore, using price promotions to improve the review label of a game appears to be a plausible strategy. I find that products that are close to upgrading their review tier are 4-9% more likely to be on sale, while products close to deteriorating their review tier are up to 4% less likely to discount, albeit this effect is less significant.

I identify two channels that can make such behavior profitable: the *selection* effect and the *variance* effect. Customers who purchase the good during a discount are likely to be different from the ones who pay the full price, and could potentially leave different reviews. If such reviews tend to be more favorable, then the induced positive selection effect creates incentives for the product to go on discount, regardless of whether the product is about to upgrade or downgrade its review tier. There is some disagreement in the literature about the sign of the selection effect. I contribute by showing that on Steam the effect is not positive.

The variance effect reflects the asymmetry between a successful and an unsuccessful discounting campaign, as measured by the valence of the new reviews. A product close to moving one review tier up benefits greatly from a good realization, while a bad might not change its review tier. On the other hand, a product that is close to sliding into a worse tier is averse to a bad outcome, and would benefit little from a good outcome. While I manage to test the selection effect directly, I have to test the variance effect by trying to falsify the predictions that follow from it. My strongest test is given by the prediction that, in the absence of a positive selection effect, the variance effect has to make a product on the verge of deteriorating its reviews less likely to discount. When I control for the size of the selection effect, I find that the aforementioned decline in probability of discounting goes from being 4% and weakly significant to 6% and highly significant. I also leverage my data and structural model to document that consumers are more likely to leave reviews when they buy during a discount, which indirectly supports the importance of the variance effect.
